### 2020-11-25 imgs/fig-lifetime.tex
Rename figure files...

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{
  arrows,
  calc,
  decorations.pathmorphing,
  decorations.pathreplacing,
  decorations.markings,
  fadings,
  positioning,
  shapes,
  arrows.meta
}
\pgfdeclareradialshading{glow}{\pgfpoint{0cm}{0cm}}{
  color(0mm)=(white);
  color(5mm)=(white);
  color(9mm)=(black);
  color(10mm)=(black)
}

\begin{tikzfadingfrompicture}[name=glow fading]
  \shade [shading=glow] (0,0) circle (1);
\end{tikzfadingfrompicture}

\ifpdf
% Ensure reproducible output
\pdfinfoomitdate=1
\pdfsuppressptexinfo=-1
\pdftrailerid{}
\fi

\begin{document}

\begin{tikzpicture}[scale=1.5391]
  \node at (0, 0) {\includegraphics[width=7.6958cm]{exp_data_norm_m_t.pdf}};
  \node at (-2.35, 1.7) {\footnotesize (\textbf{a})};

  \node at (0.77, 0.5) {$t_{\mathrm{hold}}$};
  \draw[->,>=stealth,line width=1] (0.4, 0.5) -- (0, 0.5);
  \draw[->,>=stealth,line width=1] (1.1, 0.5) -- (1.5, 0.5);
  \draw[orange!80!white,line width=1,fill=orange!50!white]
  (-0.3, 0.2) -- (-0.3, 0.8) -- node[above,orange!90!black] {\small $\pi$-pulse}
  (0, 0.8) -- (0, 0.2);
  \draw[orange!80!white,line width=1,fill=orange!50!white,shift={(1.8, 0)}]
  (-0.3, 0.2) -- (-0.3, 0.8) -- node[above,orange!90!black] {\small $\pi$-pulse}
  (0, 0.8) -- (0, 0.2);
  \draw[->,>=stealth,line width=1] (-0.5, 0.2) -- (2, 0.2);

  \node at (0, -3.9) {\includegraphics[width=7.6958cm]{exp_data_norm_a_t.pdf}};
  \node at (-2.35, 1.5 - 3.9) {\footnotesize (\textbf{b})};
  % strength = 2pi 29.3(17) mHz/mW^2.58
  \node at (0.85, -3.25) {\includegraphics[width=4.6175cm]{atomic_loss_scaling_inset.pdf}};
\end{tikzpicture}

\end{document}