### 2020-12-18 fig-theory.tex
Move figures since PRX doesn't like subdirectories

\documentclass[border={-3pt -5.5pt -5pt -4pt}]{standalone} % left bottom right top

\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{
  arrows,
  calc,
  decorations.pathmorphing,
  decorations.pathreplacing,
  decorations.markings,
  fadings,
  positioning,
  shapes,
  arrows.meta
}
\pgfdeclareradialshading{glow}{\pgfpoint{0cm}{0cm}}{
  color(0mm)=(white);
  color(5mm)=(white);
  color(9mm)=(black);
  color(10mm)=(black)
}

\begin{tikzfadingfrompicture}[name=glow fading]
  \shade [shading=glow] (0,0) circle (1);
\end{tikzfadingfrompicture}

\ifpdf
% Ensure reproducible output
\pdfinfoomitdate=1
\pdfsuppressptexinfo=-1
\pdftrailerid{}
\hypersetup{
  pdfcreator={},
  pdfproducer={}
}
\fi

\begin{document}

\begin{tikzpicture}
  \node at (-9.8, 0) {\includegraphics[width=4.5cm]{imgs/RamanTransferScheme_v6.eps}};
  \node at (-11.4, 2.2) {\footnotesize (\textbf{a})};
  \node at (-1.35, 1.25) {\includegraphics[width=11.0cm]{imgs/raman_theory_full_gamma.pdf}};
  \node at (-7.05, 2.35) {\footnotesize (\textbf{b})};
  \node at (-4.35, -1.35) {\includegraphics[width=5.7cm]{imgs/raman_v0_vhi.pdf}};
  \node at (-6.95, -0.35) {\footnotesize (\textbf{c})};
  \node at (1.4, -1.35) {\includegraphics[width=5.7cm]{imgs/wavefunction.pdf}};
  \node at (-1.2, -0.35) {\footnotesize (\textbf{d})};
\end{tikzpicture}

\end{document}